Two measures are better than one: Combining iconicity ratings and guessing experiments for a more nuanced picture of iconicity in the lexicon

### 2022-01-29 paper.Rmd
fixed up discussion

---
title: "Two measures are better than one: Combining iconicity ratings and guessing experiments for a more nuanced picture of iconicity in the lexicon"
output:
  bookdown::html_document2:
   # reference_docx: wordstyles.docx
    toc: true
    toc_float: true
    number_sections: true
    css: custom.css
    code_folding: hide
always_allow_html: true
bibliography: sources.bib

@@

```{r packages, include=FALSE}
library(tidyverse)
library(kableExtra)
library(bookdown)
library(report)
library(lmerTest)
library(plotly)
library(corrr)
library(boot)
library(zeallot)
library(pander)
library(sjPlot)
library(ggeffects)
library(gridExtra)
knitr::opts_chunk$set(echo = TRUE,warning=FALSE,message = FALSE)
```

@@

The stimuli were 304 words taken from the *Japonic Sensory Lexicon* database (ANONYMISED CITATION), which is a collection of words for 110 sensory concepts in 49 Japonic varities.

```{r}
# semantic domains of concepts, plus numbers of ideophones and non-ideophones
concepts <- read_tsv("https://raw.githubusercontent.com/BonnieMcLean/JaponicSensoryLex/main/concepts.tsv")%>%
  rowwise()%>%
  mutate(concept=str_split(concept,"-")[[1]][1])%>%
  select(concept,domain)

@@

# plus the practise and control items which arent in the Japonic sensory lexicon
words$domain[words$form == "yoroyoro"] <- "action"
words$domain[words$form == "fuwafuwa"] <- "texture"
words$domain[words$form == "hisohiso"] <- "sound"
words$domain[words$form == "pyoNpyoN"] <- "action"

```{r}
dat <- read_csv("https://raw.githubusercontent.com/BonnieMcLean/JaponicSensoryLex/main/dat.csv")
info <- 
words <- read_csv("responses-guessingwords.csv")%>%
  select(identifier)%>%
  mutate(identifier=str_replace_all(identifier,"FUZZY","CLEAR"))%>%
  unique()

cognates <- read_csv("cognatecodes.csv")%>%
  mutate(identifier=paste(test_form,eng_trans,sep="_"))%>%select(cognate,identifier)

words <- words%>%left_join(cognates,by="identifier")
```

The stimuli were 304 words from the Yamato stratum of the Japanese lexicon, in the semantic domain of sensory perception. The words were taken from the *Japonic Sensory Lexicon* (ANONYMISED CITATION), a dataset containing words in

from the domain of sensory perception, including both ideophones and prosaic lexical items. For all tasks, the words were presented to participants as audio files. The audio files were synthesised using the Google Cloud Text-to-Speech API [(https://cloud.google.com/text-to-speech)](https://cloud.google.com/text-to-speech), with the female Japanese Wavenet voice ja-JP-Wavenet-B. They were then edited using Praat [@boersma_praat_2020] to have a flat pitch of 200Hz. Expressive prosody can enhance performance in guessing tasks [@kunihira_effects_1971; @dingemanse_what_2016], and so using a flat intonation was seen as the best way to ensure comparisons between the words were fair and not influenced by the prosody in a particular recording. We also considered synthesising all the words to have a marked intonation. However, different word lengths would make changes in pitch more dramatic in shorter words compared to longer words, affecting words of different lengths unequally. While flat, synthetic voices may be the best option for comparability, they are not a natural reflection of real-world experiences with iconicity--which involve not only segments but also prosody and even multimodal cues (e.g. accompanying gesture). Thus, if anything the results in this and similar studies likely *underestimate* iconicity in language.

In all studies, participants were instructed to use headphones throughout the experiment, and could not procede to the experiment without first passing a listening test designed to require the use of headphones @woods_headphone_2017. The response buttons in each trial also did not appear until *after* the audio files had finished playing. This ensured that participants could not respond without first listening to the audio files.

@@

The words were split across 11 different experiments, in such a way that each experiment had a balance of words that were expected to be iconic, versus words that were not expected to be iconic. In addition, two 'practice words' and three 'control words' were included in every experiment. The practice items were used to give participants time to familiarise themselves with the experiment design, and were not included in the results. The control items were chosen from words found to be very iconic/highly guessable in previous studies, and were included as a check to identify participants who were perhaps not performing the task as intended. The first two trials in every experiment were the two 'practice words', while the order of the rest of the words in the experiment, including the control words, was randomly varied between participants. After finishing the experiment, participants were asked to describe the task they were performing during the experiment. This was used as an attention check, and to identify participants who had misunderstood the experiment instructions. Strange or lacking task descriptions, particularly when coupled with poor performance on the control items, were used to identify and exclude results from participants who were likely not performing the task as intended.

For the pilot version of the guessing task, in which participants had to match a given form to its correct meaning, data from 3 out of 132 participants was excluded. In one case, the participant indicated that they were choosing the translations based on whether they sounded similar to the Japanese word (rather than based on their meaning). In the other two cases, the participants did not provide a task description and performed poorly on the control items, so the data was excluded as it was not clear that they were paying attention/understood the task correctly. In the new version of the guessing task, in which participants had to match a given meaning to its correct form, data from 3 out of 332 participants was excluded. In all three cases, the participants performed poorly on the control items and either did not provide a task description, or their task descriptions were irrelevant (they commented that the task was 'interesting' or 'difficult' instead of describing what they were doing in the task).

@@

For the rating task, data from 5 out of 225 participants was excluded. For the ratings, as well as looking at task descriptions and performance on the control items, we calculated the person-total correlation for each participant @curran_methods_2016. The person-total correlation is the correlation between an individual participant's ratings, and the mean ratings from every other participant. It provides a measure of how well a participant's responses agree with those of other participants in the task. The criteria for exclusion was two or more of either (a) strange or lacking task descriptions, (b) very low ratings for control items, or (c) low person-total correlations (<.2^[We chose .2 as the cut off point because it corresponded with a discontinuity in the lower end of the distribution of person-total correlations, see Figure 1 in the online Supplementary Material]). Criteria (a) was used for 4 exclusions. In two cases, the participants indicated that they were rating the words according to how much they sounded like their English translations, rather than how well they resembled their meanings. Criteria (b) was used in 3 exclusions, and criteria (c) was used in 4 exclusions.

@@

The same formulaic structure, "After X, Y", was used for all the sentences--where Y is the target perception and X is the event that causes it. As some sensory modalities (e.g. sound) may better lend themselves to the construal of iconicity than others (e.g. taste), correct translations were always paired with foil translations from the same modality (e.g. sound translations were only paired with other sound translations, and taste translations with other taste translations, etc.). To make the difficulty of the task comparable between different trials, only words that were neither synonyms nor antonyms of the correct translation were chosen as foils. Finally, the length of the sentences was also kept consistent (to a difference of no more than 5 characters) between pairings of translations and foils. 
To test the robustness of the guessing procedure to different choices of translation and foil words, the same words were tested multiple times with different translations and foils. We were concerned that participants could choose particular translations because of how they sound (particularly if they sounded similar to the Japanese word), rather than based on their meanings as we intended. So, to minimise any potential effects from this, as much as possible phonologically distinct synonyms were chosen as alternative translations (e.g. SMALL|TINY, BIG|LARGE).

@@

## Creation of Icotools

To make the collection of guesses and ratings quicker and easier for future studies, while also improving comparability between studies, we created a python package, `icotools` [(https://pypi.org/project/icotools/)](https://pypi.org/project/icotools/), with functions that can be used to automatically generate rating and guessing experiments (using the meaning-to-word design from the second guessing experiments). The package has three functions: a rater function, a guesser function, and a foiler function. The rater and guesser functions create rating and guessing experiments as described in Sections \@ref(goodguesses) and \@ref(ratings). They take two csv files as input. The first is a wordlist containing the items in the experiment (including trials, practice items, control items, and sets of translations to use for them), while the second file is used for customisation (e.g. to specify how many items to include in each experiment, the location of media files, custom instructions and custom exit questions, etc.).

For each experiment, the functions produce three files--a html file, a php file, and a csv file. These files can then be uploaded to a server, the link to the html files can be shared with participants, and as the participants complete the experiments their responses will be written to the csv file by the php file. The functions also produce a summary csv file with all the details of the experiments (i.e. the trials in each experiment, as well as the translations and, where applicable, foils used). An R script is provided to collate the responses for all experiments into one long tidy dataset after data collection is complete.

The foiler function can be used to generate a list of opposite sounding foil words for each item in a guessing experiment, following the method described in Section \@ref(goodguesses). It is also possible to specify your own foil words, or to choose the foils randomly from among the other items in the experiment. For a full description of how to use icotools, please see the ReadMe at https://anonymous.4open.science/r/IcoTools-2CEA/README.md (this is an anonymised link to the github repository: it will later be replaced with the correct link).

@@

To test whether using phonologically distinct foils in the guessing experiments improved their sensitivity to iconicity, a subset of 30 words were tested twice--once with phonologically distinct "opposite" foils, and once with random foils.

@@

We also compared the distributions of guessing accuracies and mean iconicity ratings between the different methods. Since the iconicity ratings in this study were collected from naive participants who did not speak any Japanese, whereas rating studies more commonly use native speaker participants, I added one more plot to the figure showing the distribution of iconicity ratings in a separate study by Thompson et al. [-@thompson_iconicity_2020]. Thompson et al. collected iconicity ratings for Japanese words from native Japanese speaking participants. Their dataset contained a wider variety of words than the current dataset, and they also used a different scale. Their scale went from -5 'antiiconic' to 5 'iconic', with 0 being labelled as 'arbitrary'. The scale in this study started at 0 (also labelled as 'arbitrary', or "no resemblance between the word and its meaning"), and went up to 6 for "strong resemblance between the word and meaning". To make the data more comparable between the two studies, and because previous studies have found the negative ratings unreliable [@motamedi_iconicity_2019], I have filtered the data from Thompson et al. to only contain responses using the part of the scale between 0 and 5. I have also filtered the words to only contain verbs, adverbs, and adjectives from the Yamato stratum. The Yamato stratum refers to the native lexicon of Japanese. Thompson et al.'s study included both words from the Yamato stratum, as well as Sino-Japanese and foreign words. This study only included Yamato words, mainly verbs, adverbs and adjectives. Both the data in this study and in Thompson et al. 2020 contains a mix of ideophones and non-ideophones within the Yamato stratum. Finally, ratings from both studies were transformed so that they varied between 0 and 1, to match with the guessing accuracies. The resulting distributions are shown in Figure \@ref(fig:discriminability)

@@

everything%>%
  ggplot(aes(x=method,y=score))+geom_violin()+theme_classic()+labs(x="Method",y="Guessing accuracy / Iconicity rating")+scale_x_discrete(limits=c("guesses between translations","guesses between words","iconicity ratings","native iconicity ratings\n(Thompson et al. 2020)"))+theme(axis.text.x = element_text(angle=15,vjust=0.6))+scale_y_continuous(breaks=c(0,1))
```

Focusing first on the data from the current study, we can see that a greater spread of measures is provided by the guesses compared to the ratings, and that within the guesses, the method of guessing between words provides a better spread of measures than guessing between translations--both at the very high end of the scale (highly guessable words) and at the very low end of the scale (poorly guessable words). However, if we look at the ratings from native Japanese speakers, these have a pretty equivalent spread to the guesses between words. Their distribution is slightly narrower at the very top end of the scale. However, the data from the current study was chosen from a domain known to be high in iconicity, whereas the data from Thompson et al. was more varied, with proportionally less ideophones. Thus, this could simply be a reflection of the differences between the two datasets rather than relating to the methods used.

@@

Figure \@ref(fig:comparison) compares the guessing accuracies and mean iconicity ratings, for both ideophones and non-ideophones. The words in the pilot study where participants were guessing between translations were different to the words in the subsequent guessing and rating studies, so here only the results from the later guessing study (where participants guessed between words) are shown. At the top end of the scale, the two measures agree--ideophones score higher than non-ideophones for both measures. However, at the bottom of the scale something strange happens. While for the ratings, the bottom of the scale is dominated by non-ideophones (as expected), for the guesses, the bottom of the scale is actually dominated by *ideophones*. That is, ideophones are simultaneously guessed *better* and *worse* than non-ideophones. We will return to this in the discussion.

```{r comparison,fig.cap="Comparison of guesses and ratings, for ideophones and non-ideophones"}
# get info on ideophones
words <- read_csv("ideophone_list.csv")
words%>%
  select(form,ideophone)->info

@@

comparison_dat%>%
  filter(method=="ratings")%>%
  mutate(mean_rating=6*score)%>%
  mutate(word_type=ifelse(ideophone=="y","ideophones","non-ideophones"))%>%
  ggplot(aes(x=word_type,y=mean_rating))+geom_violin()+theme_classic()+ylim(0,6)+labs(x="Word type",y="Mean rating")->plot1

comparison_dat%>%
  filter(method=="guesses")%>%
    mutate(word_type=ifelse(ideophone=="y","ideophones","non-ideophones"))%>%
    ggplot(aes(x=word_type,y=score))+geom_violin()+theme_classic()+ylim(0,1)+labs(x="Word type",y="Guessing accuracy")->plot2

@@

Setting aside the strange distribution of guesses for the ideophones, it's notable that both measures suggest that the sensory lexicon of Japanese is for the most part *iconic*, rather than arbitrary. In the guesses, this translates to the majority of the words being guessed slightly *above*, rather than at or below chance. In the ratings, this translates to the words receiving mean ratings higher than zero (the arbitrary point on the scale)--though some of these 'iconic' ratings are likely an overstatement, since all of the words received positive iconicity ratings and yet from the guessing results we can see that many of these were not guessed any differently to chance.

To test how well the measures agree on the iconicity of individual words, iconicity ratings and guessability scores for the same words were transformed to z-scores so that they could be directly compared. Figure \@ref(fig:agreement) plots matching z-scores for the guesses and ratings against eachother, to show the agreement between the two measures. In the figure, ideophones are represented by blue dots, and non-ideophones by red dots.

```{r agreement,fig.cap="Agreement between guesses and ratings"}
comparison_dat%>%
  select(-score)%>%
  pivot_wider(names_from = "method",values_from="z_score")->comparison_dat
comparison_dat%>%
  ggplot(aes(x=guesses,y=ratings,colour=ideophone,label=identifier))+geom_point()+theme_classic()+geom_smooth(method="lm")->plot

@@

For both ideophones and non-ideophones, there was a strong correlation between the ratings and the guesses. This correspondence was slightly better for the ideophones (r = 0.66, 95% CI [0.44, 0.79], 95% CI [0.53, 0.76], t(99) = 8.68, p < .001) than for the non-ideophones (r = 0.41, 95% CI [0.29, 0.52], t(201) = 6.45, p < .001). To explore this relationship further, we created two linear regression models: one predicting guesses from ratings, ideophone status, and the interaction between these two factors, and a second model predicting ratings from guesses, ideophone status, and the interaction between these two factors. Both models again used the z-scores rather than the raw ratings and guesses. The models were created using the R package `lmerTest` (version `r packageVersion("lmerTest")`) [@lmertest].

The output of the first model (predicting guesses from ratings) is shown in Table \@ref(tab:guessesmodel).

<caption> (\#tab:guessesmodel) Linear regression model predicting guesses from ratings, ideophone status, and the interaction between these two factors. </caption>
```{r correlations and first regression}
ids <- comparison_dat%>%filter(ideophone=='y')
#report(cor.test(ids$guesses,ids$ratings))

@@

m1 <- lm(data=comparison_dat,guesses~ratings*ideophone)
#summary(m1)

@@

The model confirmed that ratings significantly predict guesses, with higher ratings leading to higher guesses. There was also a main effect of ideophone, and an interaction effect between ideophone and rating. The model is easiest to interpret by plotting the interaction, shown in Figure \@ref(fig:interaction).

```{r interaction,fig.cap="Interaction between ratings and ideophone when predicting guesses"}
ggpredict(m1,c("ratings","ideophone"))%>%plot()
```

The plot shows that higher ratings lead to better guessing accuracy for both ideophones and non-ideophones. The interaction effect appears because in the case of words with *low* iconicity ratings, guessing accuracy for ideophones is relatively poorer than for non-ideophones. At the top of the scale, highly rated ideophones also appear to be guessed slightly better than non-ideophones with comparable ratings. However, the difference here does not appear to be significant as the confidence intervals for the two lines overlap. We will return to this in the discussion.

@@

<caption> (\#tab:ratingsmodel) Linear regression model predicting ratings from guesses, ideophone status, and the interaction between these two factors. </caption>
```{r ratingsmodel,fig.cap="Model predicting guesses from ratings and ideophone status"}
m3 <- lm(data=comparison_dat,ratings~guesses*ideophone)

@@

The model shows that guesses significantly predict ratings, as does ideophone status, but that there was no interaction between these two predictors. This is shown in Figure \@ref(fig:ratingsinteraction). The lines for ideophones and non-ideophones are parallel, indicating that differences in guesses correspond to differences in ratings *in the same way* for both ideophones and non-ideophones. However, the line for the ideophones is directly *above* the line for the non-ideophones, indicating that ideophones are rated higher in iconicity than non-ideophones--even when guessed at the same accuracies.

```{r ratingsinteraction,fig.cap="Relationship between ideophone status, guessability, and predicted rating"}
ggpredict(m3,c("guesses","ideophone"))%>%plot()
```

Finally, the lower correlation between guesses and ratings obtained for non-ideophones compared to ideophones suggests that participants may not have been as consistent or reliable when rating non-ideophones compared to when rating ideophones. To further investigate this, we calculated the person-total correlation for the rating data [@curran_methods_2016; @motamedi_iconicity_2019], for ideophones and non-ideophones separately. Figure \@ref(fig:ratingconsistency) shows the agreement between raters (expressed as the correlation between each individual rating, and the by-item average) for ideophones and non-ideophones respectively. The dots indicate the mean while the lines indicate 95% confidence intervals. Again, since these ratings were from non-Japanese speaking participants, I have also added data from native Japanese speaking participants from Thompson et al. (2020) as a comparison.

@@

  # Append the results to the dataframe
  cis[nrow(cis) + 1,] = list(name, 'ideophone', posR, posLower, posUpper)
  cis[nrow(cis) + 1,] = list(name, 'non-ideophone', negR, negLower, negUpper)
}
```

```{r make person-total graph,eval=FALSE}
cis %>% 
  ggplot(aes(x=name,
             y=r,
             ymin=lower,
             ymax=upper,
             color=sign)) +
  geom_point(position=position_dodge(0.1)) +
  geom_errorbar(position=position_dodge(0.1),
                width=0.1) + 
  coord_flip() +ylim(0,0.6)+theme_classic()+labs(y="Correlation (r)",x="")->plot
plot
```

```{r ratingconsistency,fig.cap="Consistency between participants for iconicity ratings of ideophones versus non-ideophones",out.width="70%",echo=FALSE}
knitr::include_graphics("images/bootstrapping.png")
```

In both studies, participants appear to be more consistent with eachother when rating ideophones than when rating non-ideophones, though for Thompson et al. the overlap of the confidence intervals indicates that the difference was not significant. There does not appear to be any difference in the consistency of ratings *between* the English and the Japanese speakers, as both the ideophone lines and the non-ideophone lines overlap between the two groups.

@@

## Towards a reproducible workflow

>>>>

---
title: "Two measures are better than one: Combining iconicity ratings and guessing experiments for a more nuanced picture of iconicity in the lexicon"
output:
  bookdown::word_document2:
    reference_docx: wordstyles.docx
    toc: true
    toc_float: true
    number_sections: true
    css: custom.css
  #  code_folding: hide
always_allow_html: true
bibliography: sources.bib

@@

```{r setup, include=FALSE}
library(tidyverse)
library(kableExtra)
library(bookdown)
library(report)
library(lmerTest)
library(plotly)
library(corrr)
library(boot)
library(zeallot)
library(pander)
library(sjPlot)
library(ggeffects)
library(gridExtra)
library(gghalves)
knitr::opts_chunk$set(echo = FALSE,warning=FALSE,message = FALSE)
```

@@

The stimuli were 304 words taken from the *Japonic Sensory Lexicon* database [@McLean_JaponicSensoryLex_2022], which is a collection of words for 110 sensory concepts in 49 Japonic varities. The words in the database were collated from three different sources: Sanada and Tomosada [-@sanada_kenbetsu_2015; -@sanada_kenbetsu_2018], and Hirayama [-@hirayama_gendai_1992]. Any available concepts related to the senses were included. For all three sources, words for each concept were collected from speakers of different Japonic varieties by asking the speakers to translate sentences from the standard variety into their local variety. Further description of the concepts and the sentences used to elicit them can be found in the Supplementary Materials

The words used in this study come from the Yamato stratum of the Japanese lexicon, meaning that they are native Japanese words. The Yamato stratum can be further subdivided into prosaic words and ideophones [or mimetics, see e.g. @kageyama_handbook_2016]. Both types of words are found in the data from Sanada and Tomosada, whereas the data from Hirayama mainly consists of prosaic words. Figure \@ref(fig:stimuli) shows the words used in this study, including their semantic domains and membership of different lexical strata. A complete wordlist is provided in the Supplementary Materials.

```{r stimuli,fig.cap="Word used in the study, by semantic domain and lexical stratum",echo=FALSE}
# semantic domains of concepts, plus numbers of ideophones and non-ideophones
concepts <- read_tsv("https://raw.githubusercontent.com/BonnieMcLean/JaponicSensoryLex/main/concepts.tsv")%>%
  rowwise()%>%
  mutate(concept=str_split(concept,"-")[[1]][1])%>%
  select(concept,domain)

@@

# plus the practise and control items which arent in the Japonic sensory lexicon
words$domain[words$form == "yoroyoro"] <- "action"
words$domain[words$form == "fuwafuwa"] <- "texture"
words$domain[words$form == "hisohiso"] <- "sound"
words$domain[words$form == "pyoNpyoN"] <- "action"
words$domain[words$form == "zyiNzyiN"] <- "pain"
words$domain[words$form == "gaNgaN"] <- "pain"

words%>%
  unique()%>%
  group_by(domain)%>%
  count(ideophone)%>%
  mutate(stratum=ifelse(ideophone=="y","ideophone","prosaic"))->graph

graph$domain <- as.factor(graph$domain)

graph%>%
  mutate(domain=fct_relevel(domain,c("sound","action","shape & size","appearance","texture","temperature","taste","pain","bodily feeling","emotion")))%>%
  ggplot(aes(x=domain,y=n,fill=stratum))+geom_col()+theme_minimal()+labs(x="Semantic domain",fill="Lexical stratum")+theme(axis.text.x = element_text(angle=15))+scale_fill_manual(values=c("darkgrey","black"))+ coord_fixed(ratio = 0.1)->plot
```

For all tasks, the words were presented to participants as audio files. The audio files were synthesised using the Google Cloud Text-to-Speech API [(https://cloud.google.com/text-to-speech)](https://cloud.google.com/text-to-speech), with the female Japanese Wavenet voice ja-JP-Wavenet-B. They were then edited using Praat [@boersma_praat_2020] to have a flat pitch of 200Hz. Expressive prosody can enhance performance in guessing tasks [@kunihira_effects_1971; @dingemanse_what_2016], and so using a flat intonation was seen as the best way to ensure comparisons between the words were fair and not influenced by the prosody in a particular recording. We also considered synthesising all the words to have a marked intonation. However, different word lengths would make changes in pitch more dramatic in shorter words compared to longer words, affecting words of different lengths unequally. While flat, synthetic voices may be the best option for comparability, they are not a natural reflection of real-world experiences with iconicity--which involve not only segments but also prosody and even multimodal cues (e.g. accompanying gesture). Thus, if anything the results in this and similar studies likely *underestimate* iconicity in language.

In all studies, participants were instructed to use headphones throughout the experiment, and could not procede to the experiment without first passing a listening test designed to require the use of headphones [@woods_headphone_2017]. The response buttons in each trial also did not appear until *after* the audio files had finished playing. This ensured that participants could not respond without first listening to the audio files.

@@

The words were split across 11 different experiments, in such a way that each experiment had a balance of words that were expected to be iconic, versus words that were not expected to be iconic. This was so that each participant would only have to rate or guess around 34 words. In addition, two 'practice words' and three 'control words' were included in every experiment. The practice items were used to give participants time to familiarise themselves with the experiment design, and were not included in the results. The control items were chosen from words found to be very iconic/highly guessable in previous studies, and were included as a check to identify participants who were perhaps not performing the task as intended. The first two trials in every experiment were the two 'practice words', while the order of the rest of the words in the experiment, including the control words, was randomly varied between participants. After finishing the experiment, participants were asked to describe the task they were performing during the experiment. This was used as an attention check, and to identify participants who had misunderstood the experiment instructions. Strange or lacking task descriptions, particularly when coupled with poor performance on the control items, were used to identify and exclude results from participants who were likely not performing the task as intended. For the rating experiments, we also calculated the person-total correlation for each participant as an additional measure to identify participants whose response patterns differed from those of other participants in the same experiment [following @curran_methods_2016]. Based on these criteria, data from 11 out of 689 participants was excluded across the different guessing and rating tasks. Further information on these exclusions is provided in Section 1 of the Supplementary Materials.

@@

@@

The same formulaic structure, "After X, Y", was used for all the sentences--where Y is the target perception and X is the event that causes it. As some sensory modalities (e.g. sound) may better lend themselves to the construal of iconicity than others (e.g. taste), correct translations were always paired with foil translations from the same modality (e.g. sound translations were only paired with other sound translations, and taste translations with other taste translations, etc.). To make the difficulty of the task comparable between different trials, only words that were neither synonyms nor antonyms of the correct translation were chosen as foils. Finally, the length of the sentences was also kept consistent (to a difference of no more than 5 characters) between pairings of translations and foils.

To test the robustness of the guessing procedure to different choices of translation and foil words, the same words were tested multiple times with different translations and foils. We were concerned that participants could choose particular translations because of how they sound (particularly if they sounded similar to the Japanese word), rather than based on their meanings as we intended. So, to minimise any potential effects from this, as much as possible phonologically distinct synonyms were chosen as alternative translations (e.g. SMALL|TINY, BIG|LARGE).

@@

@@

To test whether using phonologically distinct foils in the guessing experiments improved sensitivity to iconicity, a subset of 30 words were tested twice--once with phonologically distinct "opposite" foils, and once with random foils.

@@

We also compared the distributions of guessing accuracies and mean iconicity ratings between the different methods. Since the iconicity ratings in this study were collected from naive participants who did not speak any Japanese, whereas rating studies more commonly use native speaker participants, we added one more plot to the figure showing the distribution of iconicity ratings in a separate study by Thompson et al. [-@thompson_iconicity_2020]. Thompson et al. collected iconicity ratings for Japanese words from native Japanese speaking participants. Their dataset contained a wider variety of words than the current dataset, and they also used a different scale. Their scale went from -5 'antiiconic' to 5 'iconic', with 0 being labelled as 'arbitrary'. The scale in this study started at 0 (also labelled as 'arbitrary', or "no resemblance between the word and its meaning"), and went up to 6 for "strong resemblance between the word and meaning". To make the data more comparable between the two studies, and because previous studies have found the negative ratings unreliable [@motamedi_iconicity_2019], we have filtered the data from Thompson et al. to only contain responses using the part of the scale between 0 and 5. We have also filtered the data to only include native Yamato words (both ideophones and prosaic words), again to increase comparability with the current study. Finally, ratings from both studies were transformed so that they varied between 0 and 1, to match with the guessing accuracies. The resulting distributions are shown in Figure \@ref(fig:discriminability)

@@

everything%>%
  ggplot(aes(x=method,y=score))+geom_half_violin()+geom_half_point()+theme_classic()+labs(x="Method",y="Guessing accuracy / Iconicity rating")+scale_x_discrete(limits=c("guesses between translations","guesses between words","iconicity ratings","native iconicity ratings\n(Thompson et al. 2020)"))+theme(axis.text.x = element_text(angle=15,vjust=0.6))+ylim(0,1)
```

Focusing first on the data from the current study, we can see that the guesses provide a more even spread of measures compared to the ratings. Between the two guessing methods, the method of guessing between words provides a wider spread of measures than guessing between translations--both at the very high end of the scale (highly guessable words) and at the very low end of the scale (poorly guessable words). If we look at the ratings from native Japanese speakers, these have a pretty equivalent spread to the guesses between words. Their distribution is slightly narrower at the very top end of the scale. However, the data from the current study was chosen from a domain known to be high in iconicity, whereas the data from Thompson et al. was more varied, with proportionally less ideophones. So, this could simply be a reflection of the differences between the two datasets rather than relating to the methods used.

@@

Figure \@ref(fig:comparison) compares the guessing accuracies and mean iconicity ratings, for both ideophones and prosaic words. The words in the pilot study where participants were guessing between translations were different to the words in the subsequent guessing and rating studies, so here only the results from the later guessing study (where participants guessed between words) are shown. At the top end of the scale, the two measures agree--ideophones score higher than prosaic words for both measures. However, at the bottom of the scale something strange happens. While for the ratings, the bottom of the scale is dominated by prosaic words (as expected), for the guesses, the bottom of the scale is actually dominated by *ideophones*. That is, ideophones are simultaneously guessed *better* and *worse* than prosaic words. We will return to this in the discussion.

```{r comparison,fig.cap="Comparison of guesses and ratings, for ideophones and prosaic words"}
# get info on ideophones
words <- read_csv("ideophone_list.csv")
words%>%
  select(form,ideophone)->info

@@

comparison_dat%>%
  filter(method=="ratings")%>%
  mutate(mean_rating=6*score)%>%
  mutate(word_type=ifelse(ideophone=="y","ideophone","prosaic"))%>%
  ggplot(aes(x=word_type,y=mean_rating))+geom_half_violin()+geom_half_point()+theme_classic()+ylim(0,6)+labs(x="Lexical stratum",y="Mean rating")->plot1

comparison_dat%>%
  filter(method=="guesses")%>%
    mutate(word_type=ifelse(ideophone=="y","ideophone","prosaic"))%>%
    ggplot(aes(x=word_type,y=score))+geom_half_violin()+geom_half_point()+theme_classic()+ylim(0,1)+labs(x="Lexical stratum",y="Guessing accuracy")->plot2

@@

Setting aside the question of the distribution of guesses for ideophones, it's notable that both measures suggest that the sensory lexicon of Japanese is for the most part *iconic*, rather than arbitrary. In the guesses, this translates to the majority of the words being guessed slightly *above*, rather than at or below chance. In the ratings, this translates to the words receiving mean ratings higher than zero (the arbitrary point on the scale). Although, some of these 'iconic' ratings are likely overstated, since all the words received positive iconicity ratings, yet from the guessing results we can see that many of these were not guessed any differently to chance.

To test how well the measures agree on the iconicity of individual words, iconicity ratings and guessability scores for the same words were transformed to z-scores so that they could be directly compared. Figure \@ref(fig:agreement) plots matching z-scores for the guesses and ratings against eachother, to show the agreement between the two measures. In the figure, ideophones are represented by blue dots, and prosaic words by red dots.

```{r agreement,fig.cap="Agreement between guesses and ratings"}
comparison_dat%>%
  select(-score)%>%
  pivot_wider(names_from = "method",values_from="z_score")->comparison_dat

comparison_dat%>%
  mutate(word_type=ifelse(ideophone=="y","ideophone","prosaic"))%>%
  ggplot(aes(x=guesses,y=ratings,colour=word_type,label=identifier))+geom_point()+theme_classic()+geom_smooth(method="lm")+labs(colour="Lexical stratum")->plot

@@

For both ideophones and prosaic words, there was a strong correlation between the ratings and the guesses. This correspondence was slightly better for ideophones (r = 0.66, 95% CI [0.44, 0.79], 95% CI [0.53, 0.76], t(99) = 8.68, p < .001) compared to prosaic words (r = 0.41, 95% CI [0.29, 0.52], t(201) = 6.45, p < .001). To explore this relationship further, we created two linear regression models: one predicting guesses from ratings, lexical stratum, and the interaction between these two factors, and a second model predicting ratings from guesses, lexical stratum, and the interaction between these two factors. Both models again used z-scores rather than raw ratings and guesses. The models were created using the R package `lmerTest` (version `r packageVersion("lmerTest")`) [@lmertest]. The output of the first model (predicting guesses from ratings) is shown in Table \@ref(tab:guessesmodel).

<caption> (\#tab:guessesmodel) Linear regression model predicting guesses from ratings, lexical stratum, and the interaction between these two factors. </caption>
```{r correlations and first regression}
ids <- comparison_dat%>%filter(ideophone=='y')
#report(cor.test(ids$guesses,ids$ratings))

@@

comparison_dat%>%
  mutate(stratum=ifelse(ideophone=="y","ideophone","prosaic"))->comparison_dat

m1 <- lm(data=comparison_dat,guesses~ratings*stratum)
#summary(m1)

@@

The model confirmed that ratings significantly predict guesses, with higher ratings leading to higher guesses. There was also a main effect of lexical stratum, and an interaction effect between stratum and rating. The model is easiest to interpret by plotting the interaction, shown in Figure \@ref(fig:interaction).

```{r interaction,fig.cap="Interaction between ratings and lexical stratum when predicting guesses"}
ggpredict(m1,c("ratings","stratum"))%>%plot()
```

The plot shows that higher ratings lead to better guessing accuracy for both ideophones and prosaic words. The interaction effect appears because for words with *low* iconicity ratings, guessing accuracy for ideophones is relatively poorer than for prosaic words. We will return to this in the discussion.

@@

<caption> (\#tab:ratingsmodel) Linear regression model predicting ratings from guesses, lexical stratum, and the interaction between these two factors. </caption>
```{r ratingsmodel,fig.cap="Model predicting guesses from ratings and ideophone status"}
m3 <- lm(data=comparison_dat,ratings~guesses*stratum)

@@

The model shows that guesses significantly predict ratings, as does lexical stratum, but that there was no interaction between these two predictors. This is shown in Figure \@ref(fig:ratingsinteraction). The lines for ideophones and prosaic words are parallel, indicating that differences in guesses correspond to differences in ratings *in the same way* for both ideophones and prosaic words. However, the line for the ideophones is directly *above* the line for prosaic words, indicating that ideophones are rated higher in iconicity than prosaic words--even when guessed at the same accuracies.

```{r ratingsinteraction,fig.cap="Relationship between ideophone status, guessability, and predicted rating"}
ggpredict(m3,c("guesses","stratum"))%>%plot()
```

Finally, the lower correlation between guesses and ratings obtained for prosaic words compared to ideophones suggests that participants may not have been as consistent or reliable when rating prosaic words compared to when rating ideophones. To further investigate this, we calculated the person-total correlation for the rating data [@curran_methods_2016; @motamedi_iconicity_2019], for ideophones and prosaic words separately. Figure \@ref(fig:ratingconsistency) shows the agreement between raters (expressed as the correlation between each individual rating, and the by-item average) for ideophones and prosaic words respectively. The dots indicate the mean while the lines indicate 95% confidence intervals. Again, since these ratings were from non-Japanese speakers, we added data from Japanese speakers from Thompson et al. (2020) as a comparison.

@@

  # Append the results to the dataframe
  cis[nrow(cis) + 1,] = list(name, 'ideophone', posR, posLower, posUpper)
  cis[nrow(cis) + 1,] = list(name, 'prosaic', negR, negLower, negUpper)
}
```

```{r make person-total graph,eval=FALSE}
cis %>% 
  ggplot(aes(x=name,
             y=r,
             ymin=lower,
             ymax=upper,
             color=sign)) +
  geom_point(position=position_dodge(0.1)) +
  geom_errorbar(position=position_dodge(0.1),
                width=0.1) +ylim(0,0.6)+theme_classic()+labs(y="Correlation (r)",x="",color="word type")->plot
plot
```

```{r ratingconsistency,fig.cap="Consistency between participants for iconicity ratings of ideophones versus prosaic words",out.width="70%",echo=FALSE}
knitr::include_graphics("images/bootstrap.png")
```

In both studies, participants appear to be more consistent with each other when rating ideophones than when rating prosaic words, though for Thompson et al. the overlap of the confidence intervals indicates that the difference was not significant. There does not appear to be any difference in the consistency of ratings *between* the English and the Japanese speakers, as both the ideophone lines and the non-ideophone lines overlap between the two groups.

@@
